Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-11-30 Response_referees.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

List of major changes apported to the paper:
\begin{itemize}
 \item We have further clarified the relation between localization properties and features of the computational basis needed
 to provide precise results. Major modifications under this perspective have been done in the introduction, in section II at pag. 5 before the analysis of fig. 2,in section IV and in the conclusions.
 \item The introduction has been deeply revised. In particular we have reduced the length of the first two paragraph concerning the differences between accuracy and precision.
 \item We added a comment after eq. (7) on the non-hermiticity of the response density and corrected the equation (11).
 \item Added a dedicated subsection (actually the IIA at page 4)  to discuss the localization properties of the FS states for static perturbations.
 \item Added a panel in figure 2 that contain the plot of the real part of $\alpha(\omega)$ in calculated in different computational domain. This new plot aims to evidence that the localization
 properties of th FS affect, on equal footing, both the real and the imaginary part of the linear response functional.
 \item We have reorganized the material after section III A. In particular we have introduced a new section (number IV in the actual version of the paper) in which we have discussed both the physical features of the excitation spectrum and the relative importance of localized vs. delocalized sector in expressing the susceptibility functional. This section also collects computational results for molecules beyond the case of $CO$ and benzene, already discussed in the previous version of the paper.
 \item We have also included in our analysis the treatment of various molecular systems, of various sizes and symmetries, such as to further illustrate the generality of our assumptions. A table has been added for comparing the excitation landscape of the molecules, in particular related to the ability of the numerical excitations to
 express the static polarizability tensor.
\end{itemize}

@@

\section*{Response to the first referee}

We wish to thank the referee for the positive review on the paper. The text has been deeply revised according to the suggestions of both the referees.

Here we report the detailed answers concerning the main aspects raised in the referee's report:
\begin{enumerate}
  \item \emph{on the point concerning the dimension of the simulation box needed to achieve the convergence of the delocalized states of the $CO$ molecule}.

  Delocalized fluctuation states do not achieve convergence as long as the simulation box is increased. Since they behave as continuum states they suffer of the same continuum collapse behavior described in section IA for the unbound solution of the Schr\"odinger equation. What is expected to converge is the spectral function, or declined in the present case the real and imaginary part of the dynamical polarizability.
  The value of convergence is system/observable and perturbation dependent.
  % It can be found through an explicit convergence analysis by introducing an acceptance tolerance, however this procedure is out of the scope of our analysis and has not been performed.

  Nonetheless, to have an idea of the dimension of the box needed to achieve convergence we can refer to the case of the Benzene molecule and make usage
  of the result found in our analysis, compared to the one reporte in fig. 2 of \cite{baroni2008}. Within our analysis we show that a box linear size of about 18 Angostroem is sufficient to produce converged results for excitation spectrum below IP (around 6.5 eV in the present case). On the contrary, results of \cite{baroni2008} evidence that to achieve convergence up to 30 eV a box of linear dimension of more than 30 Angstroem is needed, so going from the localized to the delocalized sector of excitation spectrum requires a relevant increase in the simulation domain.

@@

  \item \emph{on the point concerning the convergence rate of the delocalized vs. localized states with respect
  to the grid spacing}

 In presenting the results of this paper care has been taken to ensure the convergence of the results wrt the computational parameters. We have found that there are no relevant difference in the convergence rate of localized vs delocalized quantities (for instance excitations) in function of the grid spacing. This is due to the fact that localized and delocalized states differ in their asymptotic properties whereas the grid spacing convergence is influenced by the short-range oscillation of the functions.

@@

We answer to the critical points raised out in the report of the referee:

\begin{enumerate}

 \item \emph{On the localized sector of excitation. Dependence on the symmetry and dimension of the molecule.}

 % In section IIIA we have provided formal arguments that evidence that the excitation spectrum splits into two sectors with distinct features. The first one
 % is composed by a finite set of localized excitations with discrete eigenvalues, whereas the second one contains a continuum spectrum of delocalized objects.
 % For excitation energies lower than the ionization potential (IP) of the molecule all the excitations belong to the first sector. In the intermediate energy range which goes from IP to the absolute value of the energy of the deepest occupied orbital the two sectors compenetrate so that there are localized excitation embedded in a continuum of delocalized ones, and finally only elements of the second sector exist for higher values of the energy.
 %
 % These formal results have been supported by numerical calculations in which the excitation spectrum of several molecules has been explicitly computed.
 The attribution of each excitation of the appropriate sector passes through the evaluation of the norm of the states $\phi_p^a$ and $\chi_p^a$ that build
 the dyadic representation of the excitation (see eq. 17). Only states with norm greater than a given tolerance value are considered and the associated
 excitation is attributed to the first sector if the condition $\Omega_a < |\epsilon_p|$ with $\epsilon_a = \mathrm{max}\left(\epsilon_p\right)_{p\in a}$.
 This computational procedure provides results in optimum agreement with the formal arguments reported above. Indeed the excitation landscape of
 (for instace the panel a o fig. 5 for the case of Benzene) evidences the presence of discrete excitations (the black bullets) embedded in a pseudo-continuum
 represented by the blue filled square. Moreover the comparison between the excitation DOS computed in different computational domain shows that this way of labeling the excitation is correct since the elements attributed to the localized sector have energies that easily converge with respect to the increase of the simulation box, as expected.

### 2018-11-30 AnalyticStructure_main.tex
added futher modifications

\subsection{$CO$}
The molecule is oriented along the $z$ axis and the $z$-dimension of the box is equal to 11.5, 15.1 and 16.6 \AA.
280 unoccupied states have been computed in each setup, which span an energy range of the empty KS orbitals of 30.7, 19.3 and 15.4 eV respectively.
Such energy values should not be confused with the excitations energy ranges reported in Figs.~\ref{co_spectrum} and \ref{CO_exc}.
Once again, such choices of unoccupied states ensures convergence in of excitations spectrum in the presented range.

@@

>>>>

\subsection{$CO$}
The molecule is oriented along the $z$ axis and the $z$-dimension of the box is equal to 11.5, 15.1 and 16.6 \AA.
Bond length is 1.235\AA, stretched wrt the equilibrium LDA position such as to increase the value of the polarizability.
280 unoccupied states have been computed in each setup, which span an energy range of the empty KS orbitals of 30.7, 19.3 and 15.4 eV respectively.
Such energy values should not be confused with the excitations energy ranges reported in Figs.~\ref{co_spectrum} and \ref{CO_exc}.
Once again, such choices of unoccupied states ensures convergence in of excitations spectrum in the presented range.

@@

\subsection{Excitation Landscapes}
The attribution of each excitation of the appropriate sector passes through the evaluation of the norm of the states $\phi_p^a$ and $\chi_p^a$, in the form of Eq.~\eqref{phiandchi}. Only states with norm greater than a given tolerance value(taken as the 1\% of the total norm) are considered and the associated
excitation is attributed to the first sector if the condition $\Omega_a < |\epsilon_p|$ with $\epsilon_a = \mathrm{max}\left(\epsilon_p\right)_{p\in a}$.
This computational procedure provides results in optimum agreement with the formal arguments reported above. Indeed the excitation landscape of
(for instace the panel a of Fig.~\eqref{c6h6_excLand} for the case of Benzene) evidences the presence of discrete excitations (the black bullets) embedded in a pseudo-continuum
represented by the blue filled square. The comparison between the excitation DOS computed in different computational domain shows that this way of labeling the excitation is in line with the locality of the corresponding states since the corresponding energies easily converge with respect to the increase of the simulation domain.
